---
title: "Week4_assignment_TG"
output: html_document
date: "2024-09-19"
---


```{r}
library(tidyverse)
library(janitor)
library(rio)
```

#Load the data: https://github.com/wellsdata/CompText_Jour/raw/main/data/blackindex_master.csv

```{r}

data <- read_csv("https://raw.githubusercontent.com/wellsdata/CompText_Jour/main/data/blackindex_master.csv")
```

#Using code, describe the number of rows and columns in the dataset
```{r}
nrow(data)
ncol(data)

#the given data has 1803 rows and 30 columns
```

#Create a table that displays a count of the top 5 newspaper_city entries
```{r}
top_5 <- data %>%
count (newspaper_city) %>%
 arrange(desc(newspaper_city)) %>%
  slice_max(n, n=6)%>% filter(!is.na(newspaper_city))

```



#Create a table that counts the different entries in the "newspaper_city" column
```{r}
different_entries <- data %>%
count (newspaper_city, sort = TRUE, name = "Count") %>%
  arrange(desc(Count))
```


#Create a table that counts all newspaper_city entries by year.
#UPDATE: Filter out the NA values
```{r}
all_entries <- data %>%
count (newspaper_city, year, sort = TRUE, name = "Count") %>%
  arrange(desc(Count))%>% filter(!is.na(newspaper_city))
  

```



#Create a simple column ggplot chart that shows the total entries by year
```{r}
library(dplyr)
library(ggplot2)
  data %>%
  group_by(year) %>%
count(year) %>%
  ggplot(aes(x=year, y=n, fill=n))+
  geom_col(position = "dodge")+
labs(title = "TOTAL ENTRIES BY YEAR",
       caption = "Graphic by Teona Goderdzishvili")
  
  
```

#Armed with your data analysis, write a 250 word memo at the end of the R markdown document that summarizes your findings. Use this opportunity to ask questions or note any major flaws in the data.



This data analysis focuses on newspaper entries categorized by city and year. The dataset consists of 1,803 rows and 30 columns, covering information from 1850 to 2002.

The analysis identifies the top five cities with the highest levels of newspaper-related activities, revealing significant disparities among city entries. For example, there are 151 entries for Pittsburgh compared to just 71 for New York. This discrepancy may indicate an imbalance in newspaper production or coverage.

Additionally, the analysis highlights the different entries in the "newspaper_city" column, showing a total of 12 cities represented in the dataset.

The trend analysis by year indicates that certain years and cities have markedly more entries than others, likely influenced by significant events occurring during those times.

The column chart generated through ggplot illustrates that the peak in entries occurred in the 1920s.

However, it's important to note that a substantial portion of the data is missing. For instance, when identifying the top five newspaper_city entries, the leading entry lacked a specified city. This issue of missing information is not isolated. Addressing these gaps is crucial, not only to complete the dataset but also to understand the reasons behind the missing information and why some city entries are significantly higher than others.
